---
title: "CaseStudy_Bellabeat"
author: "Saturn.I"
date: "2023-10-01"
output: html_document
---
<br>

#### Background
*Bellabeat*, a high-tech manufacturer of health-focused products for women. *Bellabeat* is a successful small company, but they have the potential to become a larger player in the global smart device market. Urška Sršen, co-founder and Chief Creative Officer of *Bellabeat*, believes that analyzing smart device fitness data could help unlock new growth opportunities for the company.

<br>

#### Questions for the analysis
* What are some trends in smart device usage?
* How could these trends apply to *Bellabeat* customers?
* How could these trends help influence *Bellabeat* marketing strategy?

<br>

#### Business task
Generate ideas for *Bellabeat*'s products and marketing strategy.

<br>

##### Load packages
```{r}
library(tidyverse)
library(skimr)
library(ggplot2)

```

##### Import dataset
```{r}
hourly_calories <- read.csv('../Case Study 2/Fitabase Data 4.12.16-5.12.16/1_hourlyCalories_merged.csv')
hourly_intensity <- read.csv('../Case Study 2/Fitabase Data 4.12.16-5.12.16/1_hourlyIntensities_merged.csv')
hourly_steps <- read.csv('../Case Study 2/Fitabase Data 4.12.16-5.12.16/1_hourlySteps_merged.csv')

daily_activity <- read.csv('../Case Study 2/Fitabase Data 4.12.16-5.12.16/2_dailyActivity_merged.csv')
daily_weight <- read.csv('../Case Study 2/Fitabase Data 4.12.16-5.12.16/2_weightLogInfo_merged.csv')
daily_sleep <- read.csv('../Case Study 2/Fitabase Data 4.12.16-5.12.16/2_sleepDay_merged.csv')

```

##### Check datasets
Both weight and sleep data sets are not completed, so they are not further analyzed.

```{r}
hourly_calories %>% 
  group_by(Id) %>% 
  summarise(total_entries = n())

hourly_intensity %>% 
  group_by(Id) %>% 
  summarise(total_entries = n())

hourly_steps %>% 
  group_by(Id) %>% 
  summarise(total_entries = n())


daily_activity %>% 
  group_by(Id) %>% 
  summarise(total_entries = n())

daily_weight %>% 
  group_by(Id) %>% 
  summarise(total_entries = n())

daily_sleep %>% 
  group_by(Id) %>% 
  summarise(total_entries = n())

```

##### Compare hourly and daily records
Data discrepancies were found, the hourly records and daily records are not sync.

```{r}
hourly_calories %>% 
  group_by(Id) %>% 
  summarise(total_calories = sum(Calories)) %>% 
  summarise(calorie = sum(total_calories))

hourly_steps %>% 
  group_by(Id) %>% 
  summarise(total_step = sum(StepTotal)) %>% 
  summarise(steps = sum(total_step))

daily_activity %>% 
  group_by(Id) %>% 
  summarise(total_calories = sum(Calories), total_step = sum(TotalSteps)) %>% 
  summarise(calorie = sum(total_calories), steps = sum(total_step))

```

##### Combine 3 tables
```{r}
hourly_combine <- inner_join(hourly_steps, 
                             inner_join(hourly_calories, hourly_intensity, by = c('Id' = 'Id', 'ActivityHour' = 'ActivityHour')), 
                             by = c('Id' = 'Id', 'ActivityHour' = 'ActivityHour'))

```

##### Reformatting the data
```{r}
hourly_combine_2 <- hourly_combine %>% 
  mutate(id = as.character(Id),
         new_active_dtm = as.POSIXct(ActivityHour, format='%m/%d/%Y %I:%M:%S %p'), 
         new_active_date = ymd(as.POSIXct(ActivityHour, format = '%m/%d/%Y')),
         time = str_sub(new_active_dtm, -8, -7),
         .after = Id) %>% 
  select(id, new_active_dtm, new_active_date, time, StepTotal, Calories, TotalIntensity)

hourly_table <- hourly_combine_2 %>% 
  mutate(new_active_time = as.integer(hourly_combine_2$time),
         time_period = format(new_active_dtm, '%H:%M:%S'), .before = time) %>% 
  select(id, new_active_dtm, new_active_date, new_active_time, time_period, 
         total_steps = StepTotal, total_intensity = TotalIntensity, calories = Calories)


daily_table <- daily_activity %>% 
  mutate(id = as.character(Id), 
         new_active_date = as.POSIXct(ActivityDate, format = '%m/%d/%Y'),
         activity_minutes = VeryActiveMinutes+FairlyActiveMinutes+LightlyActiveMinutes) %>% 
  select(id, new_active_date, SedentaryMinutes, TotalSteps, TotalDistance, activity_minutes, Calories)

```

##### Descriptive Summary
The median of the daily walk is 7405.5 that is less than the suggest guidelines about 10,000 steps per day per adults. Source here [link](https://www.medicalnewstoday.com/articles/325323#takeaway)

The maximum sedentary time (1440) is almost triple to the maximum activity time (552).

```{r}
skim_without_charts(hourly_table)
skim_without_charts(daily_table)

```

##### Univariate Analysis
The hourly records are much skewed than the daily records.

```{r}
plot(density(hourly_table$total_intensity))
plot(density(hourly_table$total_steps))
plot(density(hourly_table$calories))

plot(density(daily_table$TotalSteps))
plot(density(daily_table$TotalDistance))
plot(density(daily_table$activity_minutes))
plot(density(daily_table$SedentaryMinutes))
plot(density(daily_table$Calories))

```

##### Binvariate Analysis
In the hourly_table, three of the factors are positively correlated.

On the other hand, Step, Distance, activity_minute and Calories are also positively correlated in the daily_table.

```{r}
pairs(hourly_table[, c("total_intensity", "total_steps", "calories")])
pairs(daily_table[, c("TotalSteps", "TotalDistance", "activity_minutes", 'SedentaryMinutes', 'Calories')])

```

##### hourly_table Plots
More steps more calories burned.

Higher intensity has more calories burned.

At first more steps lead to higher intensity, then the growth of intensity is slowed down.

```{r}
hourly_table %>%
  ggplot(aes(total_steps, calories)) +
  theme_bw() +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  labs(title = 'Steps vs Calories',
       x = 'Total Steps', y = 'Total Calories')

hourly_table %>%
  ggplot(aes(total_intensity, calories)) +
  theme_bw() +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  labs(title = 'Intensity vs Calories',
       x = 'Total Intensity', y = 'Total Calories')

hourly_table %>%
  ggplot(aes(total_steps, total_intensity)) +
  theme_bw() +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  labs(title = 'Steps vs Intensity',
       x = 'Total Steps', y = 'Total Intensity')

```

##### daily_table Plot
Steps and Distance probably in a linear relationship.

More active time has more calories burned.

On the contrary, less calories burned if the sedentary time is getting longer.

```{r}
daily_table %>% 
  ggplot(aes(TotalSteps, TotalDistance)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  labs(title = 'Steps vs Distance',
       x = 'Total Steps', y = 'Total Distance')

daily_table %>% 
  ggplot(aes(activity_minutes, Calories)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  labs(title = 'Active Minutes vs Calories',
       x = 'Active Minutes', y = 'Total Calories')

daily_table %>% 
  ggplot(aes(SedentaryMinutes, Calories)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  labs(title = 'Non Active Minutes vs Calories',
       x = 'Sedentary Minutes', y = 'Total Calories')

```

##### Further check the period with most steps
Active period for most people: 7am to 9pm.

Highly active period: 2pm to 5pm.
```{r}
hourly_table %>% 
  group_by(time_period) %>% 
  ggplot(aes(time_period, total_steps)) +
  theme_bw() +
  geom_hex() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = 'Total Steps in Hours',
       x = 'Time in Hours', y = 'Total Steps')

hourly_table %>% 
  filter(total_steps >= 6000) %>%
  distinct(time_period) %>% 
  summarise(start_time = min(time_period), end_time = max(time_period))

hourly_table %>% 
  filter(total_steps >= 9000) %>%
  distinct(time_period) %>% 
  summarise(start_time = min(time_period), end_time = max(time_period))

```

The active period records a higher intensity, and higher intensity has more calories burned.
```{r}
hourly_table %>% 
  ggplot(aes(new_active_time, total_intensity)) +
  theme_bw() +
  geom_raster(aes(fill = calories)) +
  scale_fill_gradient(low = 'white', high = 'blue') +
  xlim(0, 24) + ylim(0, 180) +
  labs(x = 'Time in Hours', y = 'Total Intensity',
       title = 'Intensity and Calories across time')

```

Though Steps have a positive correlated to calories, its effect is less strong compared to the factor of intensity.
```{r}
hourly_table %>% 
  ggplot(aes(new_active_time, total_steps)) +
  theme_bw() +
  geom_raster(aes(fill = calories)) +
  scale_fill_gradient(low = 'white', high = 'blue') +
  xlim(0, 24) + ylim(0, 180) +
  labs(x = 'Time 00-24', y = 'Total Steps',
       title = 'Steps and Calories across time')

```

I try to dive deeper into the the relationships between steps, intensity and calories.

They are positively correlated, but not sure to be in a linear relationship, so the numbers are only for reference.

The correlation of steps and calories (cor. 0.814968) is slightly lower than the intensity and calories (cor. 0.8966161).

However, the correlation of steps and intensity (cor. 0.895983) is almost the same at the intensity and calories (cor. 0.8966161).

So from the graphs and numbers, I assume steps probably one of the indicators to the intensity of activity but less likely to cause more calories burned directly; while a higher intensity of activity is more likely to have more calories burned.

```{r}
cor.test(hourly_table$total_steps, hourly_table$calories)
cor.test(hourly_table$total_intensity, hourly_table$calories)
cor.test(hourly_table$total_steps, hourly_table$total_intensity)
```


#### Summary of the analysis

**From the daily_table:-**

1. Steps, Distance, Active time and Calories are positively correlated.

2. Steps, Intensity and Calories are positvely correlated.

3. Most people have less active time and less daily walk compared to the universal guidelines.

<br>

**From the hourly_table:-**

4. Higher intensity of the activity can have more calories burned.

5. General active time for most people is between 7am to 9pm.

6. Highly active time start from 2pm to 5pm.

<br>

**To conclude the analysis:**

More steps means longer the distance, more active time and higher intensity of the activity.

The more active time and higher intensity can lead to more calories burned.

Thus, daily walking can burn more calories and have other health benefits. Source here [link](https://www.medicalnewstoday.com/articles/325323#takeaway)

<br>

Most calories burned between 7am to 9pm, especially from 2pm to 5pm.

<br>

##### Marketing Strategy
**Target Audience:**

Women who want to have work life balance.

**Key messages:** 

Raise their awareness of their daily activities. 
Encourage customers to wear *Bellabeat*'s product to achieve their healthy lifestyle goal. 

<br>

##### Recommendations on *Bellabeat*'s product:
*Bellabeat*'s product: **Bellabeat app** and **Leaf**

1. Do quality assurance on *Bellabeat*'s product for data recording.
   
   Avoid data discrepancy issue and improve data usability and accuracy.

2. Notify them to have a daily brisk walk to improve their health.
   
   Motivate people to have more activity time.
   
3. Provide default activity time period suggestion from 2pm to 5pm.
   
   People can personalize the notification in different time period to fit their schedules.

<br>

#### Acknowledgement
**Thank you** so much for the dataset from MÖBIUS [link](https://www.kaggle.com/datasets/arashnic/fitbit) and the demonstration from ANASTASIA CHEBOTINA! [link](https://www.kaggle.com/code/chebotinaa/bellabeat-case-study-with-r/notebook)

<br>

And **Thank you** for your interest to my *Bellabeat* Case Study!

This is my first time using R.
Feel free to give me any comments and recommendations for improvement!

<br>
<br>


